Followers Are Not Enough: A Multi-faceted Approach to Community Detection in Online Social Networks

### 2014-11-20 paper-postreview/methods.tex
edits

Operationally, we expect users to interact with Twitter on a human time scale, and thus the natural one-second time resolution is too fine; since most humans do not write tweets on the time scale of seconds. For this reason, we coarsen each time series by considering non-overlapping time intervals ten minutes in length. For each time interval, we record a 1 if the user has tweeted during that time interval, and a 0 if they have not. Thus, the new coarsened time series now captures whether or not the user has been active on Twitter over any given ten minute time interval in our data set. We can then compute the flow of information between two users $u$ and $v$ by computing the transfer entropy between their time series $X_{t}(u)$ and $X_{t}(v).$ See Appendix A for a detailed introduction to transfer entropy and its estimation from data.

For the communities based on transfer entropy, we weighted each edge from a user $u$ to a follower $f$ by the estimated transfer entropy of the user $u$ on $f$, 
\begin{align}
	w_{u \to f}^{\text{TE}(k)} = \widetilde{\text{TE}}_{X(u) \to X(f)}^{(k)}. \label{Eqn-EW-activity}
\end{align}
As discussed in the introduction, a positive value for the transfer entropy of the user $u$ on $f$ indicates that the time series $X(u)$ and $X(f)$, are casually related in a Granger sense, i.e, users $u$ and $f$ have PV-AC.

 We computed the transfer entropy on each coarsened time series with lag $k$ ranging from 1 to 6, this corresponds to a lag of ten minutes to an hour. The choice of lag must balance a trade-off between additional information and sparsity of samples: as we increase the lag $k$, we account for longer range dependencies, but we also decrease the number of samples available to infer a higher dimensional predictive distribution. See Appendix A for a discussion of these issues. As we will show, the underlying communities resulting from the different lags have similar structure.

>>>>

Operationally, we expect users to interact with Twitter on a human time scale, and thus the natural one-second time resolution is too fine; since most humans do not write tweets on the time scale of seconds. For this reason, we coarsen each time series by considering non-overlapping time intervals ten minutes in length. For each time interval, we record a 1 if the user has tweeted during that time interval, and a 0 if they have not. Thus, the new coarsened time series now captures whether or not the user has been active on Twitter over any given ten minute time interval in our data set. It should be noted that, as (most) users are not constantly tweeting, these time series are quite sparse. This sparsity will skew the distribution of tweet activity toward being inactive, as the distribution should be to reflect realistic interaction patterns. However, if this measure is to be used in practice a deeper analysis of the effects of binning, the resulting sparsity and their role in the activity-based community structure should be carried out.

We can then compute the flow of information from a user  $u$ to a follower $f$ by computing the transfer entropy between their time series $X_{t}(u)$ and $X_{t}(f).$ See Appendix A for a detailed introduction to transfer entropy and its estimation from data. For the communities based on transfer entropy, we weighted each edge from a user $u$ to a follower $f$ by the estimated transfer entropy of the user $u$ on $f$, 
\begin{align}
	w_{u \to f}^{\text{TE}(k)} = \widetilde{\text{TE}}_{X(u) \to X(f)}^{(k)}. \label{Eqn-EW-activity}
\end{align}
As discussed in the introduction, a positive value for the transfer entropy of the user $u$ on $f$ indicates that the time series $X(u)$ and $X(f)$, are casually related in a Granger sense, i.e, user $u$ has PV-AC with follower $f$. Said differently, there is a reduction in uncertainty about the activity of $f$ given the tweet history of user $u$.

 We computed the transfer entropy on each coarsened time series with lag $k$ ranging from 1 to 6, this corresponds to a lag of ten minutes to an hour. The choice of lag must balance a trade-off between additional information and sparsity of samples: as we increase the lag $k$, we account for longer range dependencies, but we also decrease the number of samples available to infer a higher dimensional predictive distribution.  Interestingly, as we will show later, the underlying communities resulting from the different lags have similar structure.  See Appendix A for a discussion of these issues.
